Dynamics of thin liquid film flows studied with the lattice Boltzmann method

### 2023-07-24 chapters/Paper_one.tex
Chapter two feedback Jens

\section{Introduction}
Thin layers of liquids on solid surfaces are frequently encountered in a host
of natural and technological settings \cite{de2003capillarity,Focke}. Therefore,
understanding and controlling their stability and dynamics is a central problem
for fundamental physics, as well as for applied research in process engineering
and nanotechnology~\cite{oronLongscaleEvolutionThin1997,Utada}. Coating processes, for instance,
rely crucially on the mutual affinity of liquid and surface (i.e. on
wettability properties). When the liquid film is sufficiently thin, in fact, it
can become unstable, leading to the dewetting of the coated
area~\cite{RevModPhys.81.739}. From the modelling point of view, the challenge
consists in the fact that the physics of thin films is intrinsically
multiscale, for it involves phenomena ranging from the molecular scale 
at the three phase contact line, to the micro-/nano-metric size of the 
film thickness, to the size of the film as a whole, extending over the coated substrate area.

A fully resolved bottom-up atomistic
approach would be, obviously, unfeasible, if hydrodynamic regimes are to be
explored. It clearly appears that some degree of model order reduction is required. Most hydrodynamic models of thin liquid films, in the framework of the
lubrication theory, simplify the complexity of the full 3D Navier-Stokes
equations \cite{Navier,Stokes}
to one scalar transport equation (the lubrication equation) for the film thickness field
$h(x,t)$~\cite{ReynoldsLubr,oronLongscaleEvolutionThin1997,RevModPhys.81.1131,Mitlin}:
\begin{equation}\label{eq:lubr}
    \partial_t h = \nabla \cdot \left(Q(h)\nabla p_{\mbox{\tiny{film}}}\right)
\end{equation}
Here, $Q(h)$ is the mobility function, whose explicit form depends on the
boundary condition for the velocity at the surface (for a no-slip boundary, $Q(h) = h^3/(3 \mu)$, with
$\mu$ being the dynamic viscosity), and $p_{\mbox{\tiny{film}}}$ is the film
pressure at the free liquid surface.
Stable and reliable direct numerical simulations of Eq.~(\ref{eq:lubr}) require sophisticated numerical methods, whose execution is often computationally expensive~\cite{becker2003complex}.
Moreover, an ever-growing number of microfluidic problems requires to cope with
complex fluids rather than simple liquids, i.e. fluids with non-trivial
internal microstructure and/or complex non-Newtonian rheological
behaviour (e.g. colloidal suspensions, polymer solutions, etc.). 
The quest for an efficient multiscale numerical
method for simulating thin film hydrodynamics, versatile for the inclusion of
multiphysics features, is, thus, an ongoing endeavour.

@@

When a layer of fluid is characterized by a vertical length scale $H$ much
smaller than the longitudinal one $L$, the equations of motion can be
simplified under the approximation that the ratio of the length scales, 
$\varepsilon \equiv H/L$, is small ($\varepsilon \ll1$, see Fig.~\ref{fig:scheme}).
\begin{figure}
    \centering
    \includegraphics[width = 0.75\textwidth]{graphics/Fig_1_Scheme_V3_filters_2.png}
    \caption{Schematic sketch of a model system: a thin liquid film deposited on a flat substrate. The air-liquid interface is represented by the \textit{height} $h(x,y,t)$. The characteristic thickness of the film is given by $H$.}
    \label{fig:scheme}
\end{figure}
In this limit, and for small reduced Reynolds number, $\varepsilon^2Re$
the lubrication approximation tells that the dynamics is governed by equation (\ref{eq:lubr}).
Instead of directly solving Eq.~(\ref{eq:lubr}) numerically, we follow an alternative strategy.
We build our numerical model on a class of LBMs originally proposed as solvers for
the shallow water
equations~\cite{Salmon:1999:0022-2402:503,PhysRevE.65.036309,zhou2004lattice,van2010study}.
The lattice Boltzmann equation for the discrete probability density functions of a fluid system
subject to a total force (that can include both internal and external forces)
$\mathbf{F}_{\mbox{\tiny{tot}}}$,
$f_l(\mathbf{x},t)$, reads:
\begin{equation}\label{eq:LBE}
\begin{split}
&f_l(\mathbf{x}+\mathbf{c}^{(l)}\Delta t,t+\Delta t) = \\
&(1 - \omega) f_l(\mathbf{x},t) + \omega f_l^{(eq)}(\mathbf{x},t) + w_l \frac{\Delta t}{c_s^2} \mathbf{c}^{(l)} \cdot \mathbf{F}_{\mbox{\tiny{tot}}},
\end{split}
\end{equation}
where $l$ labels the lattice velocities $\mathbf{c}_l$ and runs from $0$ to $Q-1$, with $Q$ being the number
of velocities characterizing the scheme.
Algorithmically, this equation can be seen as made up of two steps. A \textit{local collision}
step where the $f_l(\mathbf{x},t)$ ``relax'' towards the local equilibrium distributions
$f^{(eq)}_l(\mathbf{x},t)$ with rate  
$\omega = \Delta t/\tau$ (where $\tau$, the relaxation time, is proportional to the kinematic viscosity $\nu$): the distribution functions are substituted by their weighted average 
(with weights $\omega$ and $1-\omega$) 
with the equilibria, with an added so-called "source" term (the last term on the right hand side of Eq. (\ref{eq:LBE})), when a force is present. A \textit{non-local streaming} step where the updated distribution functions are scattered to the nearest neighbouring sites. 
The parameters $c_s$ (the lattice speed of sound) and $w_l$ (the 
so called "weights")
depend on the geometry of the lattice and are determined under suitable constraints on 
the form of the tensorial moments in the lattice velocities up to fourth order~\cite{wolf-gladrow}.
We work with two-dimensional square lattices of side length $N\Delta x$, with lattice constant $\Delta x$ and $Q=9$.
For simplicity, we keep $\Delta t=\Delta x=1$ throughout this paper and follow the standard notation, where $c_s = 1/\sqrt{3}$ and the $\mathbf{c}^{(l)} = (c^{(l)}_x,c^{(l)}_y)$, $l=0,1,\dots,8$, are~\cite{qian-dhumieres-lallemand,shan_yuan_chen_2006}

@@

layer at rest~\cite{PhysRevE.65.036309}.

@@

The multiscale Chapman-Enskog expansion \cite{Chapman,Enskog} of such a

@@

for the height and velocity fields~\cite{PhysRevE.65.036309,van2010study,Salmon:1999:0022-2402:503}

@@

are commonly chosen values in the literature~\cite{moulton_lega_2013,oronLongscaleEvolutionThin1997}

@@

Before concluding this section, let us notice that special care has to be taken
in the implementation of the numerical scheme, when evaluating the
forcing term since it contains higher order derivatives (the gradient $p_{\mbox{\tiny{film}}}$,
which in turn includes the Laplace
 pressure $\gamma \nabla^2 h$, see Eq.(~\ref{eq:pfilm}) and, hence,
 spurious lattice effects may arise.  
We noticed, for example, that a centered scheme to calculate gradients \cite{zhou2004lattice}
does not guarantee the sufficient degree of isotropy on the lattice as, e.g., for the relaxation of
a droplet (discussed in section~\ref{sec:results}), where it led to unphysical droplet shapes.
Therefore, we use the following expressions to compute the gradients
\begin{equation}\label{eq:derivative}
    \nabla \phi(\mathbf{x}) = 3\sum_{l=0}^8w_l\mathbf{c}^{(l)} \phi(\mathbf{x}+\mathbf{c}^{(l)})+O(\nabla^3),
\end{equation}
and the Laplacian 
\begin{align}\label{eq:n_laplace}
    \nabla^2 \phi(\mathbf{x}) =& \frac{1}{6}\bigg[4\sum_{l = \text{odd}} w_l \phi(\mathbf{x}+\mathbf{c}^{(l)})\nonumber\\ 
    &+ 1\sum_{l = \text{even}} w_l \phi(\mathbf{x}+\mathbf{c}^{(l)}) - 20\phi(\mathbf{x}) \bigg] + O(\nabla^4),
\end{align}
respectively \cite{doi:10.1137/S1064827599357188,THAMPI20131}, for a generic scalar field $\phi$ 
(be it the height field $h$, the pressure $p_{\mbox{\tiny{film}}}$ or a position dependent surface tension field).
Besides the higher degree of the isotropy, the scheme 
(\ref{eq:derivative}-\ref{eq:n_laplace}) has the advantage of employing directly the set of lattice Boltzmann speeds.

@@

Let us consider the problem of a droplet, deposited on a smooth substrate
with an apparent contact angle $\theta > \theta_{eq}$, which spreads to relax
to a shape dictated by the equilibrium contact angle $\theta_{eq}$.
The equilibrium contact angle quantifies the wettability of a given substrate by a 
certain liquid and can be calculated using Young's equation \cite{RevModPhys.57.827}
\begin{equation}\label{eq:Young}
    \gamma \cos\theta_{eq} = \gamma_{SL} - \gamma_{SG} ,
\end{equation}
with $\gamma_{SL}$ and $\gamma_{SG}$ being the surface tensions between solid/liquid and solid/gas, respectively.

@@

Let us stress that the shape is indeed very close to a spherical cap. As mentioned in 
Section \ref{sec:method}, in fact, a sufficiently accurate finite difference scheme
is required, as the one in Eqs. (\ref{eq:derivative}-\ref{eq:n_laplace})~\cite{THAMPI20131}. 
In particular, we note that the isotropy of the pressure gradient is of utmost importance: a simple scheme with two-point centered derivatives~\cite{zhou2004lattice} yields squared equilibrium droplet shapes.

>>>>

\section{Introduction}
Thin layers of liquids on solid surfaces are frequently encountered in a host
of natural and technological settings \cite{degennesCapillarityWettingPhenomena2004,Focke}. Therefore,
understanding and controlling their stability and dynamics is a central problem
for fundamental physics, as well as for applied research in process engineering
and nanotechnology~\cite{oronLongscaleEvolutionThin1997,Utada}. Coating processes, for instance,
rely crucially on the mutual affinity of liquid and surface (i.e. on
wettability properties). When the liquid film is sufficiently thin, in fact, it
can become unstable, leading to the dewetting of the coated
area~\cite{bonnWettingSpreading2009}. From the modelling point of view, the challenge
consists in the fact that the physics of thin films is intrinsically
multiscale, for it involves phenomena ranging from the molecular scale 
at the three phase contact line, to the micro-/nano-metric size of the 
film thickness, to the size of the film as a whole, extending over the coated substrate area.

A fully resolved bottom-up atomistic
approach would be, obviously, unfeasible, if hydrodynamic regimes are to be
explored. It clearly appears that some degree of model order reduction is required. Most hydrodynamic models of thin liquid films, in the framework of the
lubrication theory, simplify the complexity of the full 3D Navier-Stokes
equations \cite{Navier,Stokes}
to one scalar transport equation (the lubrication equation) for the film thickness field
$h(x,t)$~\cite{reynoldsTheoryLubricationIts1886,oronLongscaleEvolutionThin1997,crasterDynamicsStabilityThin2009,Mitlin}:
\begin{equation}\label{eq:lubr}
    \partial_t h = \nabla \cdot \left(Q(h)\nabla p_{\mbox{\tiny{film}}}\right)
\end{equation}
Here, $Q(h)$ is the mobility function, whose explicit form depends on the
boundary condition for the velocity at the surface (for a no-slip boundary, $Q(h) = h^3/(3 \mu)$, with
$\mu$ being the dynamic viscosity), and $p_{\mbox{\tiny{film}}}$ is the film
pressure at the free liquid surface.
Stable and reliable direct numerical simulations of Eq.~(\ref{eq:lubr}) require sophisticated numerical methods, whose execution is often computationally expensive~\cite{beckerComplexDewettingScenarios2003}.
Moreover, an ever-growing number of microfluidic problems requires to cope with
complex fluids rather than simple liquids, i.e. fluids with non-trivial
internal microstructure and/or complex non-Newtonian rheological
behaviour (e.g. colloidal suspensions, polymer solutions, etc.). 
The quest for an efficient multiscale numerical
method for simulating thin film hydrodynamics, versatile for the inclusion of
multiphysics features, is, thus, an ongoing endeavour.

@@

When a layer of fluid is characterized by a vertical length scale $H$ much
smaller than the longitudinal one $L$, the equations of motion can be
simplified under the approximation that the ratio of the length scales, 
$\varepsilon \equiv H/L$, is small ($\varepsilon \ll1$, see Fig.~\ref{fig:scheme}).
\begin{figure}
    \centering
    \includegraphics[width = 0.75\textwidth]{graphics/Fig_1_Scheme_V3_filters_2.png}
    \caption{Schematic sketch of a model system: a thin liquid film deposited on a flat substrate. The air-liquid interface is represented by the \textit{height} $h(x,y,t)$. The characteristic thickness of the film is given by $H$.}
    \label{fig:scheme}
\end{figure}
In this limit, and for small reduced Reynolds number, $\varepsilon^2Re$
the lubrication approximation tells that the dynamics is governed by equation (\ref{eq:lubr}).
Instead of directly solving Eq.~(\ref{eq:lubr}) numerically, we follow an alternative strategy.
We build our numerical model on a class of LBMs originally proposed as solvers for
the shallow water
equations~\cite{salmonLatticeBoltzmannMethod1999,dellarNonhydrodynamicModesPriori2002,zhouLatticeBoltzmannMethods2004,vanthangStudy1DLattice2010}.
The lattice Boltzmann equation for the discrete probability density functions of a fluid system
subject to a total force (that can include both internal and external forces)
$\mathbf{F}_{\mbox{\tiny{tot}}}$,
$f_l(\mathbf{x},t)$, reads:
\begin{equation}\label{eq:LBE}
\begin{split}
&f_l(\mathbf{x}+\mathbf{c}^{(l)}\Delta t,t+\Delta t) = \\
&(1 - \omega) f_l(\mathbf{x},t) + \omega f_l^{(eq)}(\mathbf{x},t) + w_l \frac{\Delta t}{c_s^2} \mathbf{c}^{(l)} \cdot \mathbf{F}_{\mbox{\tiny{tot}}},
\end{split}
\end{equation}
where $l$ labels the lattice velocities $\mathbf{c}_l$ and runs from $0$ to $Q-1$, with $Q$ being the number
of velocities characterizing the scheme.
Algorithmically, this equation can be seen as made up of two steps. A \textit{local collision}
step where the $f_l(\mathbf{x},t)$ ``relax'' towards the local equilibrium distributions
$f^{(eq)}_l(\mathbf{x},t)$ with rate  
$\omega = \Delta t/\tau$ (where $\tau$, the relaxation time, is proportional to the kinematic viscosity $\nu$): the distribution functions are substituted by their weighted average 
(with weights $\omega$ and $1-\omega$) 
with the equilibria, with an added so-called "source" term (the last term on the right hand side of Eq. (\ref{eq:LBE})), when a force is present. A \textit{non-local streaming} step where the updated distribution functions are scattered to the nearest neighbouring sites. 
The parameters $c_s$ (the lattice speed of sound) and $w_l$ (the 
so called "weights")
depend on the geometry of the lattice and are determined under suitable constraints on 
the form of the tensorial moments in the lattice velocities up to fourth order~\cite{wolf-gladrow}.
We work with two-dimensional square lattices of side length $N\Delta x$, with lattice constant $\Delta x$ and $Q=9$.
For simplicity, we keep $\Delta t=\Delta x=1$ throughout this paper and follow the standard notation, where $c_s = 1/\sqrt{3}$ and the $\mathbf{c}^{(l)} = (c^{(l)}_x,c^{(l)}_y)$, $l=0,1,\dots,8$, are~\cite{qian-dhumieres-lallemand,shan_yuan_chen_2006}

@@

layer at rest~\cite{dellarNonhydrodynamicModesPriori2002}.

@@

The multiscale Chapman-Enskog expansion \cite{chapmanMathematicalTheoryNonuniform1990,enskogKinetischeTheorieVorgange1917} of such a

@@

for the height and velocity fields~\cite{dellarNonhydrodynamicModesPriori2002,vanthangStudy1DLattice2010,salmonLatticeBoltzmannMethod1999}

@@

are commonly chosen values in the literature~\cite{moultonEffectDisjoiningPressure2013,oronLongscaleEvolutionThin1997}

@@

Before concluding this section, let us notice that special care has to be taken
in the implementation of the numerical scheme, when evaluating the
forcing term since it contains higher order derivatives (the gradient $p_{\mbox{\tiny{film}}}$,
which in turn includes the Laplace
 pressure $\gamma \nabla^2 h$, see Eq.(~\ref{eq:pfilm}) and, hence,
 spurious lattice effects may arise.  
We noticed, for example, that a centered scheme to calculate gradients \cite{zhouLatticeBoltzmannMethods2004}
does not guarantee the sufficient degree of isotropy on the lattice as, e.g., for the relaxation of
a droplet (discussed in section~\ref{sec:results}), where it led to unphysical droplet shapes.
Therefore, we use the following expressions to compute the gradients
\begin{equation}\label{eq:derivative}
    \nabla \phi(\mathbf{x}) = 3\sum_{l=0}^8w_l\mathbf{c}^{(l)} \phi(\mathbf{x}+\mathbf{c}^{(l)})+O(\nabla^3),
\end{equation}
and the Laplacian 
\begin{align}\label{eq:n_laplace}
    \nabla^2 \phi(\mathbf{x}) =& \frac{1}{6}\bigg[4\sum_{l = \text{odd}} w_l \phi(\mathbf{x}+\mathbf{c}^{(l)})\nonumber\\ 
    &+ 1\sum_{l = \text{even}} w_l \phi(\mathbf{x}+\mathbf{c}^{(l)}) - 20\phi(\mathbf{x}) \bigg] + O(\nabla^4),
\end{align}
respectively \cite{junkDiscretizationsIncompressibleNavier2000,thampiIsotropicDiscreteLaplacian2013}, for a generic scalar field $\phi$ 
(be it the height field $h$, the pressure $p_{\mbox{\tiny{film}}}$ or a position dependent surface tension field).
Besides the higher degree of the isotropy, the scheme 
(\ref{eq:derivative}-\ref{eq:n_laplace}) has the advantage of employing directly the set of lattice Boltzmann speeds.

@@

Let us consider the problem of a droplet, deposited on a smooth substrate
with an apparent contact angle $\theta > \theta_{eq}$, which spreads to relax
to a shape dictated by the equilibrium contact angle $\theta_{eq}$.
The equilibrium contact angle quantifies the wettability of a given substrate by a 
certain liquid and can be calculated using Young's equation \cite{degennesWettingStaticsDynamics1985}
\begin{equation}\label{eq:Young}
    \gamma \cos\theta_{eq} = \gamma_{SL} - \gamma_{SG} ,
\end{equation}
with $\gamma_{SL}$ and $\gamma_{SG}$ being the surface tensions between solid/liquid and solid/gas, respectively.

@@

Let us stress that the shape is indeed very close to a spherical cap. As mentioned in 
Section \ref{sec:method}, in fact, a sufficiently accurate finite difference scheme
is required, as the one in Eqs. (\ref{eq:derivative}-\ref{eq:n_laplace})~\cite{thampiIsotropicDiscreteLaplacian2013}. 
In particular, we note that the isotropy of the pressure gradient is of utmost importance: a simple scheme with two-point centered derivatives~\cite{zhouLatticeBoltzmannMethods2004} yields squared equilibrium droplet shapes.

### 2023-07-24 chapters/Paper_three.tex
Chapter two feedback Jens

\section{Introduction} Wet surfaces and droplets are part of our every-day experience and of numerous industrial processes including coating, tribology, painting and printing, to name but a few~\cite{gross1980fluid,szeri2010fluid,DERYCK1998278,doi:10.1146/annurev.fluid.31.1.347,dasilvasobrinhoStudyDefectsUltrathin1999,singh2010inkjet,jo2009evaluation,Wijshoff2010}. 
Moreover, the continuously growing technological interest for lab-on-a-chip devices~\cite{C6LC00387G,Focke} as well as for printable electronics~\cite{Kim_2005, Luechinger_2008} and printable photovoltaics~\cite{Brabec2001,RH20}, whose efficiency relies crucially on a precise control of material deposition upon (de-)wetting of liquid films, drew the attention to applications where the substrate is adaptive or switchable, i.e. it is not inert but responds dynamically to external stimuli or to the evolution of the coating liquid film itself~\cite{ButtEtAl_Langmuir2018,GuoGuo_RSCAdv2016}. 
Several realizations of switchable and adaptive substrates have been proposed~\cite{LiuChem2005, XiCSR2010}, involving smart materials such as polymer brushes~\cite{CohenStuartEtAl_NatMat2010,ayres2007stimuli,YongEtAl_Mat2018}, thermal-responsive hydrogels~\cite{ChenEtAl_SM2010}, light-responsive molecules and microstructures~\cite{IchimuraEtAl_Science2000,DelormeEtAl_Langmuir2005,OscuratoEtAl_AAMI2017} or processes such as electrowetting~\cite{MugeleEtAl_JPCM2005}.
Probably the simplest, yet non-trivial, modelling of a switchable substrate can be realized by a space and time dependent wettability pattern~\cite{D0SM02082F}. 
While a consistent body of theoretical/computational work has been devoted to processes on static heterogeneous substrates, though, the time dependent case is still almost unexplored, with few relevant exceptions focusing on single droplet spreading and sliding~\cite{D0SM02082F,GrawitterStark2,ThieleHartmann} or limited to an analysis of the linear regime~\cite{suman2006dynamics}.

@@

\section{Method} In order to simulate the dewetting dynamics on patterned, ``switchable'', substrates, we integrate numerically the thin-film equation~\cite{ReynoldsLubr,oronLongscaleEvolutionThin1997,RevModPhys.81.1131} 

@@

by means of a recently developed method, based on a lattice Boltzmann (LB) \\scheme~\cite{PhysRevE.100.033313,PhysRevE.104.034801, Zitz2022}.

@@

As such, it contains the information on the liquid/solid and solid/gas interactions and, hence, on the wettability, which is parameterized in terms of the contact angle $\theta$~\cite{RevModPhys.81.739, SCHWARTZ1998173}. 

@@

on the type of interactions among fluid molecules and between fluid and substrate~\cite{SCHWARTZ1998173,Mitlin,Teletzke}. A common choice, for instance, is the pair $(9,3)$, which 

@@

the dynamics remains essentially the same, but for a shift in the time scale (see also~\cite{PhysRevE.104.034801})}.

@@

A possible laboratory experiment for the here mentioned approach consists of a light switchable substrate with a thin film cast on it~\cite{becker2003complex, IchimuraEtAl_Science2000}. 

@@

Lengths and time scales will be expressed, respectively, in units of the mean film height, $h_0$ (which is constant in time, due to mass conservation), and of $t_0 = \frac{3\mu}{\gamma h_0^3 q_0^4}$, the inverse growth rate of the most unstable mode, whose wavenumber is $q_0$, of a spinodally dewetting film \cite{Mecke_2005}. On a uniform substrate, with constant contact angle $\theta^{(u)}$, the wavenumber reads 
$(q^{(u)}_0)^2 = h_{\ast}^{-2}(1-\cos \theta^{(u)})f^{\prime}(h_0/h_{\ast})$ \cite{Mecke_2005,PhysRevE.100.023108}. In our patterned case, we define 

@@

 the Hamaker constant $A_H$. Consider, for instance, the problem studied in \cite{becker2003complex}: the dewetting of a polystyrene film of thickness $h_0 \approx 4 \, \text{nm}$ deposited on a silicon 

@@

(such that $\Pi(h)= - \frac{A_H}{6\pi h^3}$)~\cite{Mecke_2005,becker2003complex}, one gets 

@@

The characteristic time of the growth of the instability for these experiments, then, reads \cite{PhysRevLett.99.114503} (the dynamic viscosity of polystyrene is 

@@

scales of the order of $t_0$ ($\approx 2 t_0$) \cite{becker2003complex}, as in our 

@@

with homogeneous substrates~\cite{PhysRevE.104.034801}).

@@

within the weak/intermediate slip regime~\cite{Peschka9275,fetzer2007quantifying, munch2005lubrication} (we have tried also a value four times smaller, without observing 

@@

The rivulet life-time can be seen as the rupture time of the structure along its axis, hence such that $\Delta h (\tau_{\text{riv}}) \sim h_0$ \cite{PhysRevE.104.034801}, which yields

@@

Considering as a reference, for instance, the system studied in ~\cite{becker2003complex,PhysRevLett.99.114503}, namely a $\sim 4 \, \text{nm}$ thick film 

>>>>

\section{Introduction} Wet surfaces and droplets are part of our every-day experience and of numerous industrial processes including coating, tribology, painting and printing, to name but a few~\cite{gross1980fluid,szeri2010fluid,DERYCK1998278,doi:10.1146/annurev.fluid.31.1.347,dasilvasobrinhoStudyDefectsUltrathin1999,singh2010inkjet,jo2009evaluation,Wijshoff2010}. 
Moreover, the continuously growing technological interest for lab-on-a-chip devices~\cite{C6LC00387G,Focke} as well as for printable electronics~\cite{Kim_2005, Luechinger_2008} and printable photovoltaics~\cite{Brabec2001,RH20}, whose efficiency relies crucially on a precise control of material deposition upon (de-)wetting of liquid films, drew the attention to applications where the substrate is adaptive or switchable, i.e. it is not inert but responds dynamically to external stimuli or to the evolution of the coating liquid film itself~\cite{ButtEtAl_Langmuir2018,GuoGuo_RSCAdv2016}. 
Several realizations of switchable and adaptive substrates have been proposed~\cite{LiuChem2005, XiCSR2010}, involving smart materials such as polymer brushes~\cite{CohenStuartEtAl_NatMat2010,ayres2007stimuli,YongEtAl_Mat2018}, thermal-responsive hydrogels~\cite{ChenEtAl_SM2010}, light-responsive molecules and microstructures~\cite{ichimuraLightDrivenMotionLiquids2000,DelormeEtAl_Langmuir2005,OscuratoEtAl_AAMI2017} or processes such as electrowetting~\cite{MugeleEtAl_JPCM2005}.
Probably the simplest, yet non-trivial, modelling of a switchable substrate can be realized by a space and time dependent wettability pattern~\cite{D0SM02082F}. 
While a consistent body of theoretical/computational work has been devoted to processes on static heterogeneous substrates, though, the time dependent case is still almost unexplored, with few relevant exceptions focusing on single droplet spreading and sliding~\cite{D0SM02082F,GrawitterStark2,ThieleHartmann} or limited to an analysis of the linear regime~\cite{suman2006dynamics}.

@@

\section{Method} In order to simulate the dewetting dynamics on patterned, ``switchable'', substrates, we integrate numerically the thin-film equation~\cite{reynoldsTheoryLubricationIts1886,oronLongscaleEvolutionThin1997,crasterDynamicsStabilityThin2009} 

@@

by means of a recently developed method, based on a lattice Boltzmann (LB) \\scheme~\cite{PhysRevE.100.033313,zitzLatticeBoltzmannSimulations2021, Zitz2022}.

@@

As such, it contains the information on the liquid/solid and solid/gas interactions and, hence, on the wettability, which is parameterized in terms of the contact angle $\theta$~\cite{bonnWettingSpreading2009, schwartzSimulationDropletMotion1998}. 

@@

on the type of interactions among fluid molecules and between fluid and substrate~\cite{schwartzSimulationDropletMotion1998,Mitlin,Teletzke}. A common choice, for instance, is the pair $(9,3)$, which 

@@

the dynamics remains essentially the same, but for a shift in the time scale (see also~\cite{zitzLatticeBoltzmannSimulations2021})}.

@@

A possible laboratory experiment for the here mentioned approach consists of a light switchable substrate with a thin film cast on it~\cite{beckerComplexDewettingScenarios2003, IchimuraEtAl_Science2000}. 

@@

Lengths and time scales will be expressed, respectively, in units of the mean film height, $h_0$ (which is constant in time, due to mass conservation), and of $t_0 = \frac{3\mu}{\gamma h_0^3 q_0^4}$, the inverse growth rate of the most unstable mode, whose wavenumber is $q_0$, of a spinodally dewetting film \cite{meckeThermalFluctuationsThin2005}. On a uniform substrate, with constant contact angle $\theta^{(u)}$, the wavenumber reads 
$(q^{(u)}_0)^2 = h_{\ast}^{-2}(1-\cos \theta^{(u)})f^{\prime}(h_0/h_{\ast})$ \cite{meckeThermalFluctuationsThin2005,zhangMolecularSimulationThin2019}. In our patterned case, we define 

@@

 the Hamaker constant $A_H$. Consider, for instance, the problem studied in \cite{beckerComplexDewettingScenarios2003}: the dewetting of a polystyrene film of thickness $h_0 \approx 4 \, \text{nm}$ deposited on a silicon 

@@

(such that $\Pi(h)= - \frac{A_H}{6\pi h^3}$)~\cite{meckeThermalFluctuationsThin2005,beckerComplexDewettingScenarios2003}, one gets 

@@

The characteristic time of the growth of the instability for these experiments, then, reads \cite{fetzerThermalNoiseInfluences2007} (the dynamic viscosity of polystyrene is 

@@

scales of the order of $t_0$ ($\approx 2 t_0$) \cite{beckerComplexDewettingScenarios2003}, as in our 

@@

with homogeneous substrates~\cite{zitzLatticeBoltzmannSimulations2021}).

@@

within the weak/intermediate slip regime~\cite{peschkaSignaturesSlipDewetting2019,fetzer2007quantifying, munchLubricationModelsSmall2005} (we have tried also a value four times smaller, without observing 

@@

The rivulet life-time can be seen as the rupture time of the structure along its axis, hence such that $\Delta h (\tau_{\text{riv}}) \sim h_0$ \cite{zitzLatticeBoltzmannSimulations2021}, which yields

@@

Considering as a reference, for instance, the system studied in ~\cite{beckerComplexDewettingScenarios2003,fetzerThermalNoiseInfluences2007}, namely a $\sim 4 \, \text{nm}$ thick film 
